\documentclass[12pt]{amsart}

\addtolength{\hoffset}{-2.25cm}
\addtolength{\textwidth}{4.5cm}
\addtolength{\voffset}{-2.5cm}
\addtolength{\textheight}{5cm}
\setlength{\parskip}{0pt}
\setlength{\parindent}{15pt}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[colorlinks = true, linkcolor = black, citecolor = black, final]{hyperref}

\usepackage{graphicx}
\usepackage{multicol}
\usepackage{ marvosym }
\usepackage{wasysym}
\newcommand{\ds}{\displaystyle}


\pagestyle{myheadings}

\setlength{\parindent}{0in}

\pagestyle{empty}

\begin{document}

\thispagestyle{empty}

{\scshape VETM 690} \hfill {\scshape \Large  Ch. 4 Inbreeding and Coancestry} \hfill {\scshape Feb 20 2021}
 
\medskip

\hrule

\bigskip

\bigskip

{\large \bf Problems}   
\begin{enumerate}
\item In a set of populations with 20 reproductive individuals each, the initial allele frequencies of a biallelic locus are $p_0 = 0.3$ and $q_0 = 0.7$. If the populations are kept under ideal conditions, what will be the expected values of the heterozygosity, the variance of allele frequencies and inbreeding at generations $t = 5, 20$ and 100?

\begin{gather*}
F_t = 1 - (1 - \frac{1}{2N})^t \\
F_5 = 1 - (1 - \frac{1}{2*20}) ^ 5 =  0.119\\
F_{20} = 1 - (1 - \frac{1}{2*20}) ^ {20} = 0.397\\  
F_{100} = 1 - (1 - \frac{1}{2*20}) ^ {100} = 0.920\\
\sigma _t^2 = p_0q_0F_t \\
\sigma_5^2 = 0.3 * 0.7 * 0.119 = 0.025\\
\sigma_{20}^2 = 0.3 * 0.7 * 0.397 = 0.083\\
\sigma_{100}^2 = 0.3 * 0.7 * 0.920 = 0.193\\
H_t = H_0(1 - \frac{1}{2N})^t \\
H_0 = 1 - \sum_{i=1}^np_i^2 = 1 - [(0.3)^2 + (0.7)^2] = 1 - 0.58 = 0.42 \\
H_5 = 0.42[1 - (\frac{1}{2*20})^5] = 0.370\\
H_{20} = 0.42[1 - (\frac{1}{2*20})^{20}] = 0.253\\
H_{100} = 0.42[1 - (\frac{1}{2*20})^{100}] = 0.033
\end{gather*}

\item In problem 4.2 we considered a plant species whose natural reproduction is a combination of autogamy and allogamy in equal proportions. From a large population of this species a line with census size $N = 20$ was founded, which was maintained for 10 generations in the greenhouse by allogamous reproduction. 
\begin{gather*}
\alpha = \frac{\beta}{2-\beta} = 0.5/(2 - 0.5) = 0.333 \\
F_{10} = 1 - (1 - \frac{1}{2N})^{10} = 1 - (1 - \frac{1}{40})^{10} = 0.224 \\
1 - F_{IT} = (1 - F_{IS})(1-F_{ST}) \\
1 - (1 - F_{IS})(1 - F_{ST}) = F_{IT} = 1 - (1 - 0.333)(1 - 0.224) = 0.482
\end{gather*}

$F_{IS}$ is inbreeding due to non-random mating in a sub-population. $F_{ST}$ is inbreeding due to population subdivision.

\item In problem 4.1 it was assumed that the locus considered is neutral. Suppose that one of the alleles of the locus is deleterious with a purging coefficient of d = 0.2. What would then be the expected values of the heterozygosity, the variance of allele frequencies and the inbreeding for this locus in generation $t = 20$?

See recursivePurgedInbreedingCoefficient.r for an mplementation of equation 4.26 which computes $g_{20}$, the purged inbreeding coefficient at generation 20, as 0.147. 

\begin{gather*}
H_0 = 0.42 \\
H_t = 2p_0q_0(1-F_t) = 2(0.3)(0.7)(1 - 0.147) = 0.358\\
\sigma^2_{q20} = p_0q_0F_{20} = (0.3)(0.7)(0.147) = 0.031
\end{gather*}

\item In a population subdivided into $n = 10$ sub-populations of constant census size N = 100, among which migration takes place with a rate $m = 0.01$ per generation following an island model, a molecular marker whose mutation rate is $u = 10^{-6}$ is analysed. (a) What is the value of the differentiation indices of $F_{ST}$ and $D$, considering or not mutation for the marker? (b) What would these values be if the marker mutation rate were $u = 10^{-3}$? 
\begin{gather*}
F_{ST}  = \frac{1}{4N((m)\frac{n}{n-1}^2 + (u)\frac{n}{n-1}) + 1} = \frac{1}{4N((0.01)\frac{10}{9}^2 + (10^{-6})\frac{10}{9}) + 1} = 0.168 \\
F_{ST} =  \frac{1}{4N((m)\frac{n}{n-1}^2) + 1} = \frac{1}{4N((0.01)\frac{10}{9}^2)+ 1} = 0.168 \\
D \approx \frac{1}{1 + \frac{m}{u(n-1)}} = \frac{1}{1 + \frac{0.01}{10^{-6}(9)}} = 0.001\\
F_{ST}  = \frac{1}{4N((m)\frac{n}{n-1}^2 + (u)\frac{n}{n-1}) + 1} = \frac{1}{4N((0.01)\frac{10}{9}^2 + (10^{-3})\frac{10}{9}) + 1} = 0.157 \\
F_{ST} =  \frac{1}{4N((m)\frac{n}{n-1}^2) + 1} = \frac{1}{4N((0.01)\frac{10}{9}^2)+ 1} = 0.168 \\
D \approx \frac{1}{1 + \frac{m}{u(n-1)}} = \frac{1}{1 + \frac{0.01}{10^{-6}(9)}} = 0.001\\
\end{gather*}


\item In an experimental fish population maintained in 10 tanks, diversity for a group of molecular markers (SNPs) is analysed obtaining an average expected heterozygosity in each tank of $H_S = 0.21$ and a total heterozygosity for the whole set of tanks $H_T = 0.32$. (a) Calculate the differentiation index $F_{ST}$, the corrected index relative to its maximum value and the $D$ index. (b) The analysis is repeated with microsatellite markers obtaining values of $H_S = 0.83$ and $H_T = 0.89$. Are the results of both types of markers consistent with each other?

\end{enumerate}

{\large \bf Self Assessment}

{\large \bf Papers}

{\large Expected vs. realized kinship of full-sibs}
\begin{itemize}
\item https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.0020041
\item Estimating additive and dominance genetic variance is based on expected proportion of genes shared by different types of relatives
\item Genome-wide coverage of markers enables the estimation of parameters like heritability using the realized degree of identity-by-descent sharing
\item this study estimated heritability of height using genome-wide identity-by-descent sharing 
\item this study showed that it is feasible to estimate genetic variance from within-family segregation, permitting the estimation of genetic variation for disease susceptibility free from confounding, non-genetic factors, also allow partitioning of genetic variation into additive and non-additive components
\item Equating observed phenotypic covariance to degree of genetic relationships from pedigree data
\item Coefficient of relationship: expected proportion of alleles that are identical-by-descent between relatives (determines additive genetic covariance between relatives)
\item Additive effects: for a given sibling pair, $\pi$ = genome-wide mean IBD is the sum of proportion shared from paternal and maternal contributions
\end{itemize}

{\large Estimated inbreeding using genomics versus pedigrees}

{\large Pedigree free genomic kinship estimates for heritability in the wild}

 {\large Genomics for quantitative genetics in the wild}

\end{document}